---
title: "Welcome to POLS 3230!"
subtitle: "Political Analysis in `R`"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include = FALSE}
options(htmltools.preserve.raw = FALSE) # https://stackoverflow.com/questions/65766516/xaringan-presentation-not-displaying-html-widgets-even-when-knitting-provided-t
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                      fig.height = 4, fig.width = 6, fig.align = 'center',
                      out.width = '100%', fig.retina = 2)
library(tidyverse) # so many useful functions!
library(here) # functions for file management
library(janitor) # functions for tidying up datasets
library(zipcodeR) # functions for mapping zip codes
library(ggthemes) # for pleasant visualization themes
library(leaflet) # for interactive maps!
library(geosphere) # functions for geographic data
library(tidytext) # functions for text analysis
library(wordcloud2) # for word clouds!
library(broom) # functions for tidying regression tables
```

```{r load data}
# load dataset
d <- read_csv(here('data/getting-to-know-you/Getting To Know You (POLS 3230) - Form Responses 1.csv')) %>% 
  # clean up variable names
  clean_names() %>%  
  # make zipcode a character for geocoding
  mutate(zipcode = as.character(hometown_zip_code_us)) %>%
  # if the zip code only has four digits, at a leading zero
  mutate(zipcode = if_else(nchar(zipcode) == 4, paste0('0', zipcode), zipcode)) %>%
  # clean up the cats/dogs and beach/mountain preference strings; anything with a "c" is cats, anything with a "b" is beaches
  mutate(cats_or_dogs = do_you_prefer_cats_or_dogs %>% 
           str_to_lower %>% 
           str_detect('c') %>% 
           if_else(. , 'Cats', 'Dogs'),
         beach_or_mountains = do_you_prefer_the_beach_or_the_mountains %>% 
           str_to_lower %>% 
           str_detect('b') %>% 
           if_else(. , 'Beach', 'Mountains'))

# geocode the zipcodes
zip_geocodes <- geocode_zip(d$zipcode)
d <- left_join(d, zip_geocodes, by = 'zipcode')

# if no zip code provided, check coordinates field
d <- d %>% 
  mutate(lng = if_else(is.na(lng), 
                       str_split(hometown_coordinates_if_you_dont_have_a_home_zip_code,
                                 pattern = ', ', simplify = TRUE)[,2] %>% 
                         as.numeric,
                       lng),
         lat = if_else(is.na(lat), 
                       str_split(hometown_coordinates_if_you_dont_have_a_home_zip_code,
                                 pattern = ', ', simplify = TRUE)[,1] %>% 
                         as.numeric,
                       lat),
         full_name = paste(first_name, last_name))

# give some variables shorter names
d <- d %>% 
  rename(height_inches = what_is_your_height_in_inches,
         comfort_with_coding = how_comfortable_are_you_with_computer_code,
         how_are_you_feeling = how_are_you_feeling_about_this_semester)
```

class: center, middle

# Where We're From

---

## Where We're From

```{r leaflet map, fig.height=6.5}
# plot everyone on a interactive leaflet map
d %>% 
  filter(!is.na(lat)) %>% 
  leaflet() %>%  
  addProviderTiles(providers$Esri.WorldGrayCanvas) %>% 
  addCircleMarkers(lng = ~lng, lat = ~lat,
                   radius = 1, color = '#BA0C2F',
                   label = ~full_name)
```

---

## Distance to Athens

```{r compute distance}
d$distance_to_athens_km <- distm(cbind(d$lng, d$lat), c(-83.36, 34) , fun = distHaversine)[,1] / 1000
```

The median student is `r d$distance_to_athens_km %>% na.omit %>% median %>% floor %>% prettyNum(big.mark = ',')` km from their hometown, but the *average* student is `r d$distance_to_athens_km %>% na.omit %>% mean %>% floor %>% prettyNum(big.mark = ',')` km from their hometown.

--

```{r distance, fig.height=5, fig.width=8}
d %>% 
  mutate(full_name = fct_reorder(full_name, -distance_to_athens_km)) %>% 
  ggplot() +
  geom_bar(mapping = aes(x=distance_to_athens_km, y = full_name,
                         fill = distance_to_athens_km),
          stat = 'identity', width = 0.4) +
  #scale_x_log10(label=scales::comma_format(accuracy = 1)) +
  scale_x_continuous(label = scales::comma_format(accuracy = 1)) +
  scale_y_discrete() +
  labs(x = 'Distance to Athens (km)', y = NULL) +
  theme_minimal() +
  guides(fill = 'none') +
  scale_fill_gradient(low = 'gray', high = '#BA0C2F')
```

---

class: center, middle

# Personal Characteristics

---

## Personal Characteristics

On average, we are `r floor(mean(d$height_inches) / 12)` feet, `r floor(mean(d$height_inches) %% 12)` inches tall.

```{r height}
# plot a histogram of heights
ggplot(data = d) +
  geom_histogram(mapping = aes(x=height_inches),
                 color = 'black') +
  theme_minimal() +
  labs(x='Height (inches)', y='Frequency')
```

---

## Personal Characteristics

```{r height and math, out.width='100%'}
ggplot(data = d) +
  geom_point(mapping = aes(x=height_inches, y=comfort_with_coding),
             alpha = 0.6) +
  facet_grid(cats_or_dogs ~ beach_or_mountains) +
  theme_bw() +
  labs(x = 'Height (inches)', y = 'Comfort With Computer Code')
```

---

## Personal Characteristics

```{r height and math 2, out.width = '70%'}
# does height correlate with math comfort?
ggplot(data = d,
       mapping = aes(x=height_inches, y=comfort_with_coding)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = 'lm', se = FALSE) +
  theme_minimal() +
  labs(x = 'Height (inches)', y = 'Comfort with Computer Code')

# model math comfort as a function of height
height_lm <- lm(comfort_with_coding ~ height_inches, data = d)
slope <- round(height_lm$coefficients['height_inches'] * 12, 2)
pval <- tidy(height_lm)$p.value[2]
```

--

`r ifelse(slope >= 0, 'Taller', 'Shorter')` people seem to be more comfortable with computer coding. On average, 1 additional foot of height is associated with `r slope` additional points in comfort. 

--

The probability that a relationship this strong could arise purely from chance is roughly `r round(pval * 100, 1)`%.

---

class: center, middle

# How We Feel

---

# How We Feel

```{r text analysis}
d_words <- d %>% 
  # create a row for each word in `how_are_you_feeling`
  unnest_tokens(word, how_are_you_feeling) %>% 
  select(full_name, word) %>% 
  # remove the stop words (i.e. boring words)
  anti_join(get_stopwords()) %>% 
  left_join(get_sentiments('bing')) %>% 
  filter(!is.na(sentiment))

# create word cloud
p1 <- d_words %>% 
  count(word) %>% 
  rename(freq = n) %>% 
  wordcloud2(color = 'random-dark', size = 0.35)

# let +1 be a positive sentiment, and -1 be a negative sentiment; get everyone's sum
people <- d_words %>% 
  mutate(sentiment = case_when(sentiment == 'positive' ~ 1,
                               sentiment == 'negative' ~ -1,
                               TRUE ~ 0)) %>% 
  group_by(full_name) %>% 
  summarize(sentiment_score = sum(sentiment))

p1
```

- `r sum(people$sentiment_score > 0)` `r if_else(sum(people$sentiment_score > 0) == 1, 'person expressed', 'people expressed')` more positive than negative feelings.
- `r sum(people$sentiment_score < 0)` `r if_else(sum(people$sentiment_score < 0) == 1, 'person expressed', 'people expressed')` more negative than positive feelings, and
- `r nrow(d) - nrow(people)` `r if_else(nrow(d) - nrow(people) == 1, 'person was', 'people were')` kinda neutral.

